---
title: "Assignment-05"
author: "Stan Xu"
date: "2026-02-25"
format:
  html:
    toc: true
    code-fold: true
  pdf:
    documentclass: article
bibliography: references.bib
---

## 2

```{r}
#| label: load-packages
library(knitr)
library(tidyverse)
wdi <- read.csv('wdi.csv')
```

## 3

```{r}
#| label: eda-summary
#| echo: true

# Summary of selected indicators
wdi %>%
  select(gdp_per_capita, inflation_rate, life_expectancy) %>%
  summary()
```

## 3,4

##### GDP growth rate


## 4

#### Relationship between GDP per Capita and Life Expectancy

```{r}
#| label: fig-gdp-life
#| fig-cap: "Relationship between GDP per Capita and Life Expectancy.(Source:WDI, 2022)"
#| echo: false

ggplot(wdi, aes(x = gdp_per_capita, y = life_expectancy)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  theme_minimal() +
  labs(
    title = "GDP per Capita vs. Life Expectancy (2022)",
    x = "GDP per Capita (USD)",
    y = "Life Expectancy (Years)"
  )
```
Findings: The scatter plot shows a clear positive relationship between GDP per capita and life expectancy in 2022. Countries with higher GDP per capita generally have higher life expectancy. The upward-sloping regression line indicates that as income increases, life expectancy tends to increase as well. However, the relationship appears to flatten at higher income levels

#### Education Expenditure (% GDP)

```{r}
#| label: fig-gdp
#| fig-cap: "Distribution of GDP Growth Rate (Source:WDI, 2022)"

ggplot(wdi, aes(x = gdp_growth_rate)) +
  geom_histogram(fill = "steelblue", bins = 20) +
  labs(
    title = "Distribution of GDP Growth Rate",
    x = "GDP Growth Rate (%)",
    y = "Count"
  ) +
  theme_minimal()
```

The boxplot compares education expenditure as a share of GDP between high-growth and low-growth countries. It shows that both groups allocate a similar proportion of GDP to education, with medians around 3--5%.

Findings: The histogram of GDP growth rates shows that most countries cluster around modest positive growth between approximately –5% and 10%. However, the distribution has small number of countries experiencing extremely high growth rates (above 20% and even near 60%) and a few cases of severe economic contraction (around –30%).



#### 5
```{r}
#| label: tbl-key-stats
#| tbl-cap: "Average Global Indicators for 2022"

wdi_summary <- wdi %>%
  summarise(
    Avg_GDP_Capita = mean(gdp_per_capita, na.rm = TRUE),
    Avg_GDP_Growth = mean(gdp_growth_rate, na.rm = TRUE),
    Avg_Inflation = mean(inflation_rate, na.rm = TRUE),
    Avg_Life_Exp = mean(life_expectancy, na.rm = TRUE)
  )

kable(wdi_summary, digits = 2)
```


#### 7
The data for this analysis is sourced from the @worldbank2022. The second research by @rosling2011 shows that the historical gap between wealth and health outcomes is gradually closing as shown in @fig-gdp-life.






