---
title: "Assignment05"
author: "Stan Xu"
format: 
  revealjs:
    theme: sky
    transition: slide
---

## Introduction
* This presentation covers a sample of the 2022 World Development Indicators dataset.


## Loading packages and data

```{r}
#| label: load-packages
#| echo: true
library(knitr)
library(tidyverse)
wdi <- read.csv('wdi.csv')
```

## Explortary Data Analysis

```{r}
#| label: eda-summary
#| echo: true

wdi %>%
  select(gdp_per_capita, inflation_rate, life_expectancy) %>%
  summary()
```

## Wealth vs. Health

```{r}
#| label: fig-gdp-life
#| fig-cap: "Relationship between GDP per Capita and Life Expectancy.(Source:WDI, 2022)"
#| echo: false

ggplot(wdi, aes(x = gdp_per_capita, y = life_expectancy)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  theme_minimal() +
  labs(
    title = "GDP per Capita vs. Life Expectancy (2022)",
    x = "GDP per Capita (USD)",
    y = "Life Expectancy (Years)"
  )
```

Findings: The scatter plot shows a clear positive relationship between GDP per capita and life expectancy in 2022. Countries with higher GDP per capita generally have higher life expectancy. The upward-sloping regression line indicates that as income increases, life expectancy tends to increase as well. However, the relationship appears to flatten at higher income levels


## Distribution of GDP Growth Rate

```{r}
#| label: fig-gdp
#| fig-cap: "Distribution of GDP Growth Rate (Source:WDI, 2022)"

ggplot(wdi, aes(x = gdp_growth_rate)) +
  geom_histogram(fill = "steelblue", bins = 20) +
  labs(
    title = "Distribution of GDP Growth Rate",
    x = "GDP Growth Rate (%)",
    y = "Count"
  ) +
  theme_minimal()
```


Findings: The histogram of GDP growth rates shows that most countries cluster around modest positive growth between approximately --5% and 10%. However, the distribution has small number of countries experiencing extremely high growth rates (above 20% and even near 60%) and a few cases of severe economic contraction (around --30%).

## Key Variable table

```{r}
#| label: tbl-key-stats
#| tbl-cap: "Average Global Indicators for 2022"

wdi_summary <- wdi %>%
  summarise(
    Avg_GDP_Capita = mean(gdp_per_capita, na.rm = TRUE),
    Avg_GDP_Growth = mean(gdp_growth_rate, na.rm = TRUE),
    Avg_Inflation = mean(inflation_rate, na.rm = TRUE),
    Avg_Life_Exp = mean(life_expectancy, na.rm = TRUE)
  )

kable(wdi_summary, digits = 2)
```


## Citation

The data for this analysis is sourced from the @worldbank2022. The second research by @rosling2011 shows that the historical gap between wealth and health outcomes is gradually closing as shown in @fig-gdp-life.



## Conclusion

Economic Impact: Countries with higher GDP per capita generally show better health outcomes, as seen in the positive trend of life expectancy.

Economic Stability: While the vast majority of the world experienced typical, modest economic growth in 2022, the overall global picture includes extreme outliers.


Thank you!